---
title: "HW Class 11 Pt.2 (Population analysis) [Extra Credit BoxPlot]"
author: "Jimmi"
format: pdf
---

# Population Analysis

>Q13: Read this file into R and determine the sample size for each genotype and their corresponding median expression levels for each of these genotypes.

Let's read data and store it into a dataframe.

```{r}
gene <- read.table(file='rs8067378_ENSG00000172057.6.txt', header=TRUE)
head(gene)
```

Identifying how many samples have a specific genotype.

```{r}
genotypes = table(gene$geno)
genotypes
```

We can find useful data with the `summary()` function and specifically get the median by using the built in R `median()` function.

```{r}
summary( gene$exp[gene$geno == "A/A"] )
summary( gene$exp[gene$geno == "A/G"] )
summary( gene$exp[gene$geno == "G/G"] )

median(gene$exp[gene$geno == "A/A"])
median(gene$exp[gene$geno == "A/G"])
median(gene$exp[gene$geno == "G/G"])
```

The sample size for A/A is `r genotypes[1]` with an `r median(gene$exp[gene$geno == "A/A"])` median expression level.

The sample size for A/G is `r genotypes[2]` with an `r median(gene$exp[gene$geno == "A/G"])` median expression level.

The sample size for G/G is `r genotypes[3]` with an `r median(gene$exp[gene$geno == "G/G"])` median expression level.

>Q14: Generate a boxplot with a box per genotype, what could you infer from the relative expression value between A/A and G/G displayed in this plot? Does the SNP effect the expression of ORMDL3?

Turning genotypes are factors before loading up ggplot to make a boxplot.

```{r}
gene$geno = as.factor(gene$geno)
head(gene)
```

```{r}
library(ggplot2)

ggplot(gene, aes(x=geno, y=exp,fill=geno)) + 
  geom_boxplot(notch=TRUE) +
  geom_jitter(alpha=0.25, fill="black",width = 0.2) + 
  theme(legend.position="right",
        plot.title = element_text(size=11)) +
  labs(x="Genotype", y = "Expression")
```

We can infer that A/A is incurs the most expression of ORMDL3 and dominant while G/G seen in underexpressed ORMDL3 and recessive. The SNP loci appears to have an effect on the ORMDL3 gene, however, there are obvious variable ranges between each genotypes. Therefor, without a statistical significance test it is not certain.



