---
title: "hw6 function"
author: "Jimmi Nguyen"
date: "2023-01-30"
output: pdf_document
---

> Q6. How would you generalize the original code above to work with any set of input protein structures?

must load bio3d package using `install.packages("bio3d")` then call it using the `library()` function.

```{r}
library(bio3d)
  #Input to function is the character name of the protein of interest
ProteinPlot = function(protein){
  #will read and obtain protein from pdb database
  x = read.pdb(protein)
  #obtains and stores subset of atoms in Alpha chain from protein
  x.chainA = trim.pdb(x, chain="A", elety="CA")
  #obtains and stores b atoms from protein
  x.b = x.chainA$atom$b
  #plots protein line graphs with residues versus b-factor
  plotb3(x.b, sse=x.chainA, typ="l", ylab="Bfactor")
}
```

To use function, call it using `ProteinPlot()` and inputing the protein name as the input argument such as **"4AKE"**, **"1AKE"**, and **"1E4Y"**.
`ProteinPlot()` will output a line plot of the protein with with residues versus b-factor

```{r}
ProteinPlot("4AKE")
ProteinPlot("1AKE")
ProteinPlot("1E4Y")
```

